---
title: "DataMiningProject"
author: "Hemanth Kumar Kottapalli, SriNikitha Desetty, Trilok Dhankekula"
date: "2024-04-06"
output: html_document
---

# Loading dataset
```{r}


library(readxl)
library(ggplot2)
library(DataExplorer)
library(caret)



dataset_retail <- read_excel("C:/Users/shiva/Downloads/Dataset 13-Online Retail.xlsx")
str(dataset_retail)
head(dataset_retail)

```
# Number of variables and obseravtions in the data set
```{r}
observations <- nrow(dataset_retail)  # Calculate the number of observations
variables <- ncol(dataset_retail)     # Calculate the number of variables
print(paste("Number of observations:", observations))  # Print the number of observations
print(paste("Number of variables:", variables))

```

# Data Exploration
```{r}

dataset_retail$InvoiceNo <- as.factor(dataset_retail$InvoiceNo)     # Convert InvoiceNo to factor
dataset_retail$StockCode <- as.factor(dataset_retail$StockCode)     # Convert StockCode to factor
dataset_retail$CustomerID <- as.factor(dataset_retail$CustomerID)   # Convert CustomerID to factor
dataset_retail$InvoiceDate <- as.Date(dataset_retail$InvoiceDate, format = "%m/%d/%y")  # Convert InvoiceDate to Date format
str(dataset_retail)  # Display the structure of the modified dataset
```

# Install and load the stringr package for string manipulation
```{r}

# install.packages("stringr")
library(stringr)

# Change description to title case using str_to_title() function from stringr package
dataset_retail$Description <- str_to_title(dataset_retail$Description)

# Replace "United Kingdom" with "UK" in the Country column using ifelse() function
# if the value in the Country column is "United Kingdom", replace it with "UK", otherwise keep the original value
dataset_retail$Country <- ifelse(dataset_retail$Country == "United Kingdom", "UK", dataset_retail$Country)

# Display the first few rows of the dataset
head(dataset_retail)

```

## Handling Missing Values
```{r}

# Calculate the total number of missing observations in the dataset
missing_values <- sum(is.na(dataset_retail))
cat("Missing Observations are", missing_values)

# Create Data Frame for Missing Values
Variables = c("InvoiceNo", "StockCode", "Description", "Quantity", "InvoiceDate", "UnitPrice", "CustomerID", "Country")
NA_Count=c(sum(is.na(dataset_retail$InvoiceNo)),sum(is.na(dataset_retail$StockCode)),sum(is.na(dataset_retail$Description)),sum(is.na(dataset_retail$Quantity)),sum(is.na(dataset_retail$InvoiceDate)),sum(is.na(dataset_retail$UnitPrice)),sum(is.na(dataset_retail$CustomerID)),sum(is.na(dataset_retail$Country)))
df_missing_values <- data.frame(Variables,NA_Count)
df_missing_values
```


```{r}

# Remove rows with missing values and create a new dataset
dataset_update <- na.omit(dataset_retail)

# Calculate the total number of missing values in the updated dataset
sum_missing <- sum(is.na(dataset_update))

# Display the total number of missing values
print(paste("Total number of missing values after removal:", sum_missing))

# Display the dimensions of the updated dataset
print(paste("Dimensions of the updated dataset:", dim(dataset_update)))


```

Following the removal of null observations, the dataset now comprises 406,829 entries across 8 variables. Further exploration using visualization methods can help identify potential outliers.

## Visualization


```{r}
# Create a boxplot for the 'UnitPrice' variable in the updated dataset
boxplot_Price <- boxplot(dataset_update$UnitPrice,
                          main = "Price Box Plot",
                          ylab = "Unit Price",
                          notch = TRUE)
```

The UnitPrice boxplot reveals an extremely high outlier, indicating a potential source of bias in the mean and median values of the data. Omitting this outlier is essential to prevent skewed results and ensure the accuracy of statistical measures.


```{r}

# Create a boxplot for the 'Quantity' variable in the updated dataset
boxplot_Quantity <- boxplot(dataset_update$Quantity,
                             main = "Quantity Box Plot",
                             ylab = "Sales",
                             notch = FALSE)

```

The Quantity boxplot highlights outliers significantly exceeding the median value, indicating potential data entry errors or bulk purchases. The skewed distribution suggests most orders consist of smaller quantities. Addressing these outliers through data transformation or cleaning is crucial for accurate predictive analysis, ensuring more reliable models devoid of biases.



```{r}

# Calculate the frequency of each country in the updated dataset
frequency <- table(dataset_update$Country)
frequency <- sort(frequency, decreasing = TRUE)

# Select the top 5 countries with the highest frequencies
top_frequency <- head(frequency, 5)

# Create a bar plot to visualize the frequency of the top 5 countries
barplot(top_frequency,
        main = "Country Box Plot",
        xlab = "Frequency",
        ylab = "Country",
        las = 1,
        col = "green")


```
The bar plot illustrates the frequency of buyers across various countries, with the number of frequent buyers depicted on the x-axis and participating countries on the y-axis. Notably, the United Kingdom, the location of the store with a substantial presence, has been excluded to prevent biased results. Among the remaining countries, Germany emerges as the second-highest in terms of buyer frequency, closely followed by france and EIRE This visualization offers insights into the distribution of buyers across different regions, highlighting significant markets beyond the UK.


## calculate and filter the data based on the "Quantity" variable and use the interquartile range (IQR) method to identify and remove outliers.

```{r}

dataset_update$Sales <- dataset_update$UnitPrice * dataset_update$Quantity
# Display dimensions after calculating total sales
dim(dataset_update)

library(dplyr)
# Calculate the interquartile range (IQR) for Quantity
Sales_IQR <- IQR(dataset_update$Sales, na.rm = TRUE)

# Calculate the first quartile (Q1) and third quartile (Q3) for Quantity
Sales_Q1 <- quantile(dataset_update$Sales, 0.25, na.rm = TRUE)
Sales_Q3 <- quantile(dataset_update$Sales, 0.75, na.rm = TRUE)

# Calculate the lower and upper bounds for outliers
Sales_LB <- Sales_Q1 - 3 * Sales_IQR
Sales_UB <- Sales_Q3 + 3 * Sales_IQR

# Filter the dataset to include only observations where Quantity falls within the calculated bounds
filter_dataset <- dataset_update[dataset_update$Sales >= Sales_LB & dataset_update$Sales <= Sales_UB,]
# Display dimensions after filtering by sales bounds
dim(filter_dataset)

# Filter the dataset to exclude observations where UnitPrice is less than or equal to 0.01
filter_dataset <- filter_dataset %>% filter(UnitPrice >= 0.01)
# Display dimensions after further filtering by unit price
dim(filter_dataset)

```

This code snippet first calculates the total sales by multiplying the unit price and quantity columns and then filters the dataset to include only observations where the total sales fall within a specified range. Finally, it further filters the dataset to exclude observations where the unit price is less than or equal to 0.01, ensuring data integrity for subsequent analyses.

```{r}
# Creates a new variable 'UKCountryGroup' based on 'Country', labeling 'UK' or 'Other'.

filter_dataset$UKCountryGroup <- ifelse(filter_dataset$Country == "UK", "UK", "Other")

# Defines a function to determine seasons from dates: Spring, Summer, Fall, Winter.
func_season <- function(date) {
  month <- as.numeric(format(date, "%m"))
  if (month %in% 3:5) {
    return("Spring")
  } else if (month %in% 6:8) {
    return("Summer")
  } else if (month %in% 9:11) {
    return("Fall")
  } else {
    return("Winter")
  }
}
# Applies 'func_season' to create 'Season' column based on 'InvoiceDate'.
filter_dataset$Season <- sapply(filter_dataset$InvoiceDate, func_season)

table(filter_dataset$Season)

```
Creates 'UKCountryGroup' variable based on 'Country', classifying 'UK' or 'Other', and assigns seasons to 'InvoiceDate' observations, generating 'Season' column.
```{r}
# Display dimensions of the filtered dataset
dim(filter_dataset)

# Calculate the number of observations and variables in the filtered dataset
no_of_observations <- nrow(filter_dataset)
no_of_variables <- ncol(filter_dataset)

# Print the number of observations and variables
cat("There are ", no_of_observations,"observations and ", no_of_variables, "variables in the dataset after handling outliers")
```


```{r}
filter_dataset$InvoiceDummy <- ifelse(substr(filter_dataset$InvoiceNo, 1, 1) == 'C', 1, 0)

table(filter_dataset$InvoiceDummy)
```
Creates a new binary variable 'InvoiceDummy' indicating whether the invoice starts with 'C' (1 for True, 0 for False), and then displays the frequency table of 'InvoiceDummy'.


```{r}
## Dividing the dataset into Retail Sales and Retail return sales
online_sales <- filter_dataset %>% filter(InvoiceDummy == 0)

head(online_sales)
```

```{r}

# Retail return sales
online_return <- filter_dataset %>%filter(InvoiceDummy == 1)

head(online_return)
```
```{r}
nooutliers_Quantity_Boxplot <- boxplot(online_sales$Quantity,main="Quantity BoxPlot", ylab="Sales",col="green",notch=FALSE)
```
```{r}
Price_boxplot_nooutliers <- boxplot(online_sales$UnitPrice,main = "Unit Price BoxPlot",ylab = "Unit Price",col = "blue",notch = TRUE)
```

```{r}
online_sales$Log_Sales <- log(online_sales$Sales)  # Compute the natural logarithm of 'Quantity' and store it in 'Log_Sales'
online_sales$Log_UnitPrice <- log(online_sales$UnitPrice)  # Compute the natural logarithm of 'UnitPrice' and store it in 'Log_UnitPrice'

# Display the first few rows of the dataset with logarithmically transformed sales and unit price
head(online_sales)


```

```{r}
boxplot_sales_nooutliers <- boxplot(online_sales$Log_Sales,
main = "use of log transformation for Quantity Box Plot",ylab = "Sales",col = "red",notch = TRUE)
```
```{r}
boxplot_UnitPrice_nooutliers <- boxplot(online_sales$Log_UnitPrice,main = "use of log transformation for UnitPrice Box Plot",ylab = "Unit Price",col = "blue",notch = TRUE)
```

```{r}
# Calculate the frequency of each country in the online_sales dataset
frequency_country <- table(online_sales$Country)

# Sort the frequencies in decreasing order
frequency_country <- sort(frequency_country, decreasing = TRUE)

# Select the top 5 countries with the highest frequencies
top_country_frequency <- head(frequency_country, 5)

# Create a horizontal bar plot to visualize the frequency of the top 5 countries
barplot(top_country_frequency,
  main = "Bar Plot of Country (After handling outliers)",
  xlab = "Frequency",
  ylab = "Country",
  horiz = TRUE,  # Horizontal orientation
  las = 1,       # Rotate labels perpendicular to the axis
  cex.names = 0.8,  # Adjust the size of the country names
  col = "purple"  # Color of the bars
)
```

```{r}

# Aggregate total quantities by invoice number
total_invoice_numbers <- aggregate(Quantity ~ InvoiceNo, data = online_sales, FUN = sum)

# Sort invoices by total quantity in descending order
total_invoice_numbers <- total_invoice_numbers[order(-total_invoice_numbers$Quantity), ]

# Select the top 30 invoices for visualization
no_of_invoices <- 30
top_invoices <- head(total_invoice_numbers, no_of_invoices)


# Assuming top_invoices is correctly prepared
ggplot(top_invoices, aes(x = reorder(InvoiceNo, Quantity), y = Quantity)) +
  geom_bar(stat = "identity", fill = "green") +
  theme_minimal() +
  labs(title = "Top Invoices by Quantity",
       x = "Invoice No",
       y = "Quantity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

```{r}

# Aggregate total quantities by customer ID
customer_total_quantity <- aggregate(Quantity ~ CustomerID, data = online_sales, FUN = sum)

# Sort customers by total quantity in descending order
customer_total_quantity <- customer_total_quantity[order(-customer_total_quantity$Quantity), ]

# Select the top 20 customers for visualization
no_of_customers <- 20
top_20 <- head(customer_total_quantity, no_of_customers)

# Create a bar chart to display top customers by quantity
ggplot(top_20, aes(x = reorder(as.factor(CustomerID), Quantity), y = Quantity)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  theme_minimal() +
  labs(title = paste("Top 20 Customers Quantity by Customer ID"),
       x = "Customer ID",
       y = "Total Quantity") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))

```

```{r}

# Calculate total sales returns by customer
returns_sale <- online_return %>%group_by(CustomerID) %>%summarise(Sales = sum(Sales)) %>%arrange(Sales)

# Select the top 10 customers with the highest sales returns
top_returns <- head(returns_sale, 10)


# Create a bar plot to visualize total sales returns by customer
ggplot(top_returns, aes(x = reorder(CustomerID, -Sales), y = Sales, fill = Sales)) +
  geom_bar(stat = "identity") +coord_flip() +scale_fill_gradient(low = "orange", high = "grey") +labs(title = "Total Sales Returns by Customer",subtitle = "Negative values represent the amount returned",x = "Customer ID",y = "Total Sales Returns") +theme_minimal() +geom_text(aes(label = Sales), hjust = -0.1, size = 3) + theme(axis.text.y = element_text(angle = 0))




```
```{r}

# Load the 'lubridate' package
library(lubridate)

# Extract the day of the week from the InvoiceDate column and store it in a new column
filter_dataset$DayOfWeek <- wday(filter_dataset$InvoiceDate, label = TRUE, abbr = FALSE)

# Aggregate sales by day of the week
sales_pattern <- filter_dataset %>%
  group_by(DayOfWeek) %>%
  summarize(Sales = sum(Sales, na.rm = TRUE)) %>%
  mutate(DayOfWeek = factor(DayOfWeek, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")))

# Create a bar plot to visualize the sales pattern per day of the week
ggplot(sales_pattern, aes(x = DayOfWeek, y = Sales, fill = DayOfWeek)) +
  geom_col() +
  labs(title = "Weekly Sales Pattern",
       x = "Day",
       y = "Sales") +
  scale_y_continuous(labels = scales::comma) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r}

# Aggregate total sales by customer ID
summary_of_customer <- aggregate(Sales ~ CustomerID, data = online_sales, FUN = sum)

# Create a correlation matrix with selected variables from the filtered dataset
corr_matrix <- online_sales[, c('UnitPrice', 'Quantity', 'Sales')]

# Calculate the correlation matrix
corr_matrix_unitPrice_Quantity <- cor(corr_matrix)

# Display the correlation matrix as a table
knitr::kable(corr_matrix_unitPrice_Quantity)
```
```{r}

# Load required libraries
library(dplyr)
library(tidyr)

# Group data by product description and calculate total sales for each product
pivot_sales <- online_sales %>%group_by(Description) %>%summarise(Sales = sum(Sales, na.rm = TRUE)) %>%
  ungroup() %>%arrange(desc(Sales))

# Display the top rows of the pivot_sales dataset
head(pivot_sales)


```

```{r}
# Group data by country and calculate total sales for each country
pivot_country <- online_sales %>%group_by(Country) %>%
  summarise(Sales = sum(Sales, na.rm = TRUE)) %>%ungroup() %>%arrange(desc(Sales))

# Print the pivot_country dataset
print(pivot_country)

```

```{r}

# Aggregate total sales by country
total_sales_by_country <- aggregate(Sales ~ Country, data = online_sales, sum)

# Order countries by total sales in descending order
total_sales_by_country <- total_sales_by_country[order(-total_sales_by_country$Sales), ]

# Select the top 5 countries by sales
top_5<- head(total_sales_by_country, 5)

# Create a bar plot for the top 5 countries
barplot(
  top_5$Sales,names.arg = top_5$Country,
  las = 2,
  main = "Top 5 Countries Sales",
  col = "green",
  cex.names = 0.7,
  ylim = c(0, max(top_5$Sales) * 1.1)
)

```


```{r}

# Calculate the frequency of the dependent variable
frequency_sales <- table(online_sales$Sales)

# Display the first few entries of the frequency table
head(frequency_sales)
```

```{r}

# Create a bar plot to visualize the frequency distribution of sales
barplot(frequency_sales,
        main = "Distribution of Sales",
        xlab = "Sales",
        ylab = "Frequency")
```



```{r}

# Calculate the median of total sales
total_sales_median <- median(online_sales$Sales, na.rm = TRUE)

# Create a new column 'TopSale' based on whether the total sales are above or below the median
online_sales$TopSale <- ifelse(online_sales$Sales > total_sales_median, '0', '1')

# Display the frequency of each category in the 'TopSale' column
table(online_sales$TopSale)

```

```{r}

# Calculate the 75th percentile of Quantity and 25th percentile of UnitPrice
percentile_75forQuantity <- quantile(online_sales$Quantity, 0.75)
percentile_25forUnitPrice <- quantile(online_sales$UnitPrice, 0.25)

# Create a new column 'Segments' based on the conditions
online_sales$Segments <- with(online_sales, ifelse(Quantity >= percentile_75forQuantity & UnitPrice <= percentile_25forUnitPrice, 'High_Quantity',ifelse(Quantity < percentile_75forQuantity & UnitPrice > percentile_25forUnitPrice, 'Low_Quantity', 'Other')))

High_Quantity_Low_Price_set <- subset(online_sales, Segments == 'High_Quantity')
Low_Quantity_High_Price_set <- subset(online_sales, Segments == 'Low_Quantity')

# Descriptive statistics
summary(High_Quantity_Low_Price_set)
summary(Low_Quantity_High_Price_set)

# Visualization
# Scatter plot for High Quantity and Low Price
plot(High_Quantity_Low_Price_set$Quantity, High_Quantity_Low_Price_set$UnitPrice, main="High Quantity and Low Price", xlab="Quantity", ylab="Unit Price", col="orange")

# Scatter plot for Low Quantity and High Price
plot(Low_Quantity_High_Price_set$Quantity, Low_Quantity_High_Price_set$UnitPrice, main="Low Quantity and High Price", xlab="Quantity", ylab="Unit Price", col="green")

# Assuming 'Segments' is a variable in your 'online_sales' dataset
sales_pattern_segment <- online_sales %>%
  group_by(Segments) %>%
  summarise(Sales = sum(Sales))

# Print the table
print(sales_pattern_segment)

```

```{r}

# Calculate correlation between Quantity and UnitPrice for High Quantity, Low Price segment
correlation_High_Quantity_Low_Price <- cor(High_Quantity_Low_Price_set$Quantity, High_Quantity_Low_Price_set$UnitPrice)

# Calculate correlation between Quantity and UnitPrice for Low Quantity, High Price segment
correlation_Low_Quantity_High_Price <- cor(Low_Quantity_High_Price_set$Quantity, Low_Quantity_High_Price_set$UnitPrice)

# Print correlation results
print(correlation_High_Quantity_Low_Price)
print(correlation_Low_Quantity_High_Price)

# Perform linear regression for High Quantity, Low Price segment
model_High_Quantity_Low_Price <- lm(UnitPrice ~ Quantity, data = High_Quantity_Low_Price_set)

# Perform linear regression for Low Quantity, High Price segment
model_Low_Quantity_High_Price <- lm(UnitPrice ~ Quantity, data = Low_Quantity_High_Price_set)

# Summarize regression results for High Quantity, Low Price segment
summary(model_High_Quantity_Low_Price)

# Summarize regression results for Low Quantity, High Price segment
summary(model_Low_Quantity_High_Price)


```



```{r}


# Load the caret package
library(caret)
library(lattice)

# Set the random seed for reproducibility
set.seed(123)

# Sample 60% of the data for training
no_of_obs_train <- sample(rownames(online_sales), dim(online_sales)[1] * 0.6)
train_data <- online_sales[no_of_obs_train, ]

# Select the remaining data for validation
no_of_obs_test <- setdiff(rownames(online_sales), no_of_obs_train)
test_data <- online_sales[no_of_obs_test, ]

# Print the dimensions of the training and validation sets
dim(train_data)
dim(test_data)

```




```{r}

library(dplyr)
# Calculate order frequency for each customer
order_frequency_customers <- online_sales %>%group_by(CustomerID) %>%
  summarize(OrderFrequency = n_distinct(InvoiceNo))

# Determine quantiles for order frequency
quantiles_order <- quantile(order_frequency_customers$OrderFrequency, probs = c(0.25, 0.75))

# Categorize customers based on order frequency
order_frequency_customers$Frequency.customer <- cut(order_frequency_customers$OrderFrequency,
                                          breaks = c(-Inf, quantiles_order, Inf),
                                          labels = c("Rare", "Moderate", "Frequent"),
                                          include.lowest = TRUE)

# Merge customer frequency information into the main dataset
online_sales <- merge(online_sales, order_frequency_customers, by = "CustomerID")


```




```{r}

names(train_data)

# Use only selected predictors based on domain knowledge or prior analysis
mlr_model <- lm(Log_Sales ~ Season + UKCountryGroup, data = train_data)
summary(mlr_model)



```



# Multiple Linear regression with all predictors (ID's excluded)


```{r}
## Load necessary libraries
library(caret)

# Assuming train_data and test_data are already loaded and contain the appropriate columns

# Set the random seed for reproducibility
set.seed(123)

# Defining the model excluding identified arbitrary variables
model_formula <- Log_Sales ~ Quantity + InvoiceDate  + Country + 
                 UKCountryGroup + Season + InvoiceDummy + Log_UnitPrice + TopSale 

# Fit the model on the training data
model <- lm(model_formula, data = train_data)

# Summary of the model to view coefficients and statistics
summary(model)

# Testing the model on the test set
test_predictions <- predict(model, newdata = test_data)

# Comparing actual values vs predicted values
results <- data.frame(Actual = test_data$Log_Sales, Predicted = test_predictions)

# Calculate RMSE to evaluate the model
rmse <- sqrt(mean((results$Actual - results$Predicted)^2))
rmse

# Displaying the RMSE
print(rmse)

```

```{r}
## Load necessary libraries
library(caret)

# Assuming train_data and test_data are already loaded and contain the appropriate columns

# Set the random seed for reproducibility
set.seed(123)

# Defining the model excluding identified arbitrary variables
model_formula2 <- Log_Sales ~   UKCountryGroup + Season + Log_UnitPrice+TopSale 

# Fit the model on the training data
model2 <- lm(model_formula2, data = train_data)

# Summary of the model to view coefficients and statistics
summary(model2)

# Testing the model on the test set
test_predictions2 <- predict(model2, newdata = test_data)

# Comparing actual values vs predicted values
results2 <- data.frame(Actual = test_data$Log_Sales, Predicted = test_predictions2)

# Calculate RMSE to evaluate the model
rmse2 <- sqrt(mean((results$Actual - results2$Predicted)^2))
rmse2

# Displaying the RMSE
print(rmse2)
```










# Decision Tree model


```{r}

library(rpart.plot)
library(rpart)

# Set the random seed for reproducibility
set.seed(123)

# Defining the target variable and predictors
# Define the model formula
# Define the model formula
model_formula <- Log_Sales ~   UKCountryGroup + Season + Log_UnitPrice+TopSale

# Fit the decision tree model on the training data
dt_model <- rpart(model_formula, data = train_data, method = "anova")

# Plot the decision tree (optional)
rpart.plot(dt_model, main = "Decision Tree for Log_Sales")



```




```{r}
# Predict on the test data for a regression tree
predictions <- predict(dt_model, newdata = test_data)

# Actual values from the test data (assuming 'Log_Sales' is your target variable)
actual_values <- test_data$Log_Sales

# Calculate RMSE to evaluate the model
rmse <- sqrt(mean((predictions - actual_values)^2))

# Print RMSE
print(rmse)


```




```{r}

# Assuming train_data and test_data are loaded and model formulas are correctly defined
# Fit the Linear Regression Model
lm_model <- lm(Log_Sales ~ Quantity + InvoiceDate + UnitPrice + Country + Sales + 
               UKCountryGroup + Season + InvoiceDummy + Log_UnitPrice + TopSale + Segments, 
               data = train_data)
# Predict with Linear Regression Model
lm_predictions <- predict(lm_model, newdata = test_data)

# Fit the Decision Tree Model
dt_model <- rpart(Log_Sales ~ Quantity + InvoiceDate + UnitPrice + Country + Sales +
                  UKCountryGroup + Season + InvoiceDummy + Log_UnitPrice + TopSale + Segments, 
                  data = train_data, method = "anova")
# Predict with Decision Tree Model
dt_predictions <- predict(dt_model, newdata = test_data)

# Calculate RMSE for Linear Regression Model
lm_rmse <- sqrt(mean((test_data$Log_Sales - lm_predictions)^2))
print(paste("RMSE for Multiple Linear Regression Model:", lm_rmse))

# Calculate RMSE for Decision Tree Model
dt_rmse <- sqrt(mean((test_data$Log_Sales - dt_predictions)^2))
print(paste("RMSE for Decision Tree Model:", dt_rmse))

```






```{r}
library(caret)
library(rpart.plot)

# Set the random seed for reproducibility
set.seed(123)

# Ensure train_data is a dataframe and inspect first few rows
if (!is.data.frame(train_data)) {
  stop("train_data needs to be a dataframe")
}
print(head(train_data))

# Correct the model formula
model_formula <- Log_Sales ~   UKCountryGroup + Season + Log_UnitPrice+TopSale

# Define the control function for training with cross-validation
fitControl <- trainControl(
  method = "cv",
  number = 10,
  savePredictions = "final",
  returnResamp = "all",
  verboseIter = TRUE
)

# Train the model
dt_model_cv <- train(model_formula, data = train_data, method = "rpart",
                     trControl = fitControl,
                     tuneLength = 10)

# Plot the best tree
rpart.plot(dt_model_cv$finalModel, main = "Optimized Decision Tree for Log_Sales")

# Predict on the test data
test_predictions <- predict(dt_model_cv, newdata = test_data)

# Calculate RMSE for test data
test_rmse <- sqrt(mean((test_data$Log_Sales - test_predictions)^2))
print(paste("Test RMSE:", test_rmse))


```


Model: Random Forest Model

```{r, warning = FALSE, message = FALSE}
## Running Random Forest model


library(randomForest)

set.seed(123)
sales_rf1 <- randomForest(Log_Sales ~   UKCountryGroup + Season + Log_UnitPrice+TopSale, data = train_data, ntree = 500, importance = TRUE)

print(sales_rf1)

```





```{r}
rf_prediction1 <- predict(sales_rf1, newdata = test_data)

# Calculating RMSE
rmse3 <- sqrt(mean((rf_prediction1 - test_data$Log_Sales)^2))

# Printing the RMSE
print(rmse3)

rf_prediction2 <- predict(sales_rf1, newdata = test_data)

# Calculating RMSE
rmse4 <- sqrt(mean((rf_prediction2 - test_data$Log_Sales)^2))

# Printing the RMSE
print(rmse4)

```



